Transition path sampling based calculations of free energies of enzymatic reactions: the case of human MAT2A and plasmodium vivax adenosine deaminase

### 2022-04-14 tex/manuscript.tex
added some more text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  Transition path sampling (TPS) has been widely applied for the
  calculations of transition state and reaction coordinates of 
  condensed phase systems. Here, we discuss a scheme for the calculation
  of free energies with the TPS method 
  We discuss a scheme for the calculation of free energies within 
  the transition path sampling (TPS) method. TPS is often used for
  the calculations of reaction coordinates of complex reactions in 
  the condensed phase and in this work we use it to sample the 
  equilibrium distribution of the reaction coordinate. This 
  distribution can then be used for the calculations of free energies
  using standard Boltzmann inversion schemes. We calculate the free
  energy profiles of the reactions catalyzed by the human methionine 
  S-adenosyltransferase (MAT2A) enzyme and the plasmodium 
  vivax adenosine deaminase enzyme to demonstrate the applicability of this 
  method. 
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Molecular dynamics (MD) simulations are increasingly being
used to calculate quantitative estimations of experimentally measurable 
kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
\cite{Karplus02NatStructMolBiol9p646,
Schramm98AnnuRevBiochem67p693,Zhang05AccChemRes38p379,
Schramm11AnnuRevBiochem80p703,Schramm18ChemRev118p11194} 
This has helped in proposing mechanisms and explanations for 
enzyme activity, fast high throughput screening of thousands of inhibitor 
molecues in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
directed evolution for designing enzymes with enhanced catalytic 
activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
Straightforward MD simulations adequately sample 
the long-lived stable reactant and product states 
but the high energy barrier-crossing events that occur rarely and at long 
time scales are not easily accessible using this method alone.
To access such states at a higher frequency using MD simulations 
several enhanced sampling techniques 
have been developed in the past few decades some of which are 
umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
steered MD, \cite{Park04JChemPhys120p5946}
milestoning, \cite{Faradjian04JChemPhys120p10880} and transition path sampling 
(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}

@@

As order parameters that define the reactant and product basins, we
used the donor-hydride and acceptor-hydride distances. The
reactant basin was defined as a donor-hydride distance smaller
than 1.6  and the product basin as an acceptor-hydride
distance smaller than 1.6 . To seed the generation of the path
ensemble, we generated a biased trajectory by applying a
harmonic force between the hydride and the acceptor and
allowed the system to propagate for 250 fs. New trajectories
were generated by the shooting algorithm: we perturbed the
momenta of a randomly chosen slice of a reactive trajectory,
and using the new momenta, we allowed the system to
propagate forward and backward in time for 250 fs. A total of
120 reactive trajectories were generated.

\subsubsection{plasmodium vivax ADA}
%------------------------------------------------------------------------------ 
\subsection{Free energy calculations}
%------------------------------------------------------------------------------ 
For the free energy calculations we divided the reaction coordinate into 13 windows for 
MAT2A and 20 windows for pvADA with a $15\%$ overlap on each side of the window.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  Transition path sampling (TPS) has been widely applied for the
  calculations of reaction rate, transition state and reaction 
  coordinates of condensed phase systems. Here, we discuss a scheme 
  for the calculation of free energies with the TPS method in combination 
  with a window based sampling technique. This 
  distribution can then be used for the calculations of free energies
  using standard Boltzmann inversion schemes. We calculate the free
  energy profiles of the reactions catalyzed by the human methionine 
  S-adenosyltransferase (MAT2A) enzyme and the plasmodium 
  vivax adenosine deaminase enzyme to demonstrate the applicability of this 
  method. 
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Molecular dynamics (MD) simulations are increasingly being
used to calculate quantitative estimations of experimentally measurable 
kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
\cite{Karplus02NatStructMolBiol9p646,
Schramm98AnnuRevBiochem67p693,Zhang05AccChemRes38p379,
Schramm11AnnuRevBiochem80p703,Schramm18ChemRev118p11194} 
This has helped in proposing mechanisms and explanations for 
enzyme activity, fast high throughput screening of thousands of inhibitor 
molecues in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
directed evolution for designing enzymes with enhanced catalytic 
activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
Straightforward MD simulations adequately sample 
the long-lived stable reactant and product states 
but the high energy barrier-crossing events that occur rarely and at long 
time scales are not easily accessible using this method alone.
To access such states at a higher frequency using MD simulations 
several enhanced sampling techniques 
have been developed in the past few decades some of which are 
umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
steered MD, \cite{Park04JChemPhys120p5946}
milestoning, \cite{Faradjian04JChemPhys120p10880} and transition path sampling 
(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}

@@

The key bond parameters for the reaction in this system consists of the 
oxygen atom on the ATP, the carbon atom on the adenosyl group and the 
sulphur atom on the L-methionine. The reactant basin is defined by the 
oxygen-carbon bond distance being less than 1.7 {\AA} and the product basin 
is characterized by the suphur-carbon bond distance being less than 
1.7 {\AA}. The initial reactive trajectory was generated by the application 
of a bias harmonic force between the oxygen and the carbon atoms 
and running dynamics for $250\;fs$. The shooting algorithm was used to generate
new trajectories within which first a random time slice is chosen from the reactive 
trajectory and perturbing the momenta of the system. Then with the new momenta
the system is propagated forward and backward in time for $250\;fs$.
A total of 120 reactive trajectories were generated.

\subsubsection{plasmodium vivax ADA}
The $pv$ADA reaction is 
%------------------------------------------------------------------------------ 
\subsection{Free energy calculations}
%------------------------------------------------------------------------------ 
For the free energy calculations we divided the reaction coordinate into 13 windows for 
MAT2A and 20 windows for pvADA with a $15\%$ overlap on each side of the window.
